\texttt{yacoub}: a Python package for Simulating Generalized Fading Channels

### 2018-03-14 paper/yacoub.tex
working on majorization-minimization

\begin{abstract}
    We present a well tested Python-based library for simulating and computing
    generalized fading channels, named \texttt{yacoub}. We describe the
    applicability of \texttt{yacoub} using examples in recent communications
    systems challenges, namely: cooperative spectrum sensing, bit error rate computation
    in generalized fading channel, and parameter estimation in free space optics.
    The development of \texttt{yacoub} open source and its code is avaliable at
    \texttt{http://github.com/mirca/yacoub}.
\end{abstract}

@@

\subsection{Note on notation}
Scalars and random variables are denoted as \textit{italic} small-case letters \textit{e.g.} $x$;
vectors and random vectors are denoted as \textit{italic}, boldface, small-case letters \textit{e.g.} $\bm{x}$.
The $n-$th component of a vector $\bm{x}$ is denoted as $x_n$.
A complex vector of length $n$ is defined as $\bm{x} \in \mathbb{C}^{n\times 1}$. All
vectors are column vectors. Matrices are denoted as \textit{italic}, boldface, capital
letters as in $\bm{X}$; the identity matrix of order $n$ is denoted as $\bm{I}_n$.
We define a discrete-time circularly symmetric Gaussian process $\bm{z}$ as any (finite or infinite) collection
of random varibles $\bm{z}~=~\bm{x}~+~j\bm{y}$, $j \triangleq \sqrt{-1}$, such that
$\bm{x}$ and $\bm{y}$ are iid jointly Gaussian with zero mean vector and covariance
matrix given by $\mathbb{E}\left[\bm{z}\bm{z}^{\dagger}\right]$, in which $\bm{z}^\dagger$
means the conjugate transpose of $\bm{z}$ and $\mathbb{E}$ denotes the expected value. The
probability of an event $A$ is denoted as $\mathbb{P}(A)$.

@@

The pdf of $\tilde{y}$ can be written using the Law of Total Expectation
\begin{align}
    p(\tilde{y} | H_1) = \mathbb{E}\left[p(\tilde{y} | h, H_{1})\right]
                 = \int_{-\infty}^{+\infty} p(\tilde{y} | h, H_1)p(h)\;\mathrm{d}h.
\end{align}

Recall that the energy detection rule can be expressed as
\begin{equation}
    d_\delta (\tilde{y}) =
\left\{
    \begin{array}{ll}
        1, & \tilde{y} \geq \delta,\\
        0, & \tilde{y} < \delta,
    \end{array}
\right.
\end{equation}
in which $\delta$ is a strictly positive real number known as energy threshold,
and $d_\delta (\tilde{y}) = j,~j \in \{0,1\}$ means that the detector has decided
in favor of the hypothesis $H_j$.

As a result, the probabilities of false alarm and detection can
be written as
\begin{align}
    p_f = 1 - p()\\
    p_d =
\end{align}

\subsection{Parameter Estimation in Free Space Optics}

>>>>

\begin{abstract}
    We present a well tested Python-based library for simulating and computing
    generalized fading channels, named \texttt{yacoub}. We describe the
    applicability of \texttt{yacoub} using examples in recent communications
    systems challenges, namely: spectrum sensing, bit error rate computation,
    and parameter estimation in generalized fading channels. For the latter,
    a Majorization-Minimization algorithm is derived. The development of
    \texttt{yacoub} open source and its code is avaliable at
    \texttt{http://github.com/mirca/yacoub}.
\end{abstract}

@@

\subsection{Note on notation}
Scalars and random variables are denoted as \textit{italic} small-case letters \textit{e.g.} $x$;
vectors and random vectors are denoted as \textit{italic}, boldface, small-case letters \textit{e.g.} $\bm{x}$.
The $n-$th component of a vector $\bm{x}$ is denoted as $x_n$.
A complex vector of length $n$ is defined as $\bm{x} \in \mathbb{C}^{n\times 1}$. All
vectors are column vectors. Matrices are denoted as \textit{italic}, boldface, capital
letters as in $\bm{X}$; the identity matrix of order $n$ is denoted as $\bm{I}_n$.
We define a discrete-time circularly symmetric Gaussian process $\bm{z}$ as any (finite or infinite) collection
of random varibles $\bm{z}~=~\bm{x}~+~j\bm{y}$, $j \triangleq \sqrt{-1}$, such that
$\bm{x}$ and $\bm{y}$ are iid jointly Gaussian with zero mean vector and covariance
matrix given by $\mathbb{E}\left[\bm{z}\bm{z}^{\dagger}\right]$, in which $\bm{z}^\dagger$
means the conjugate transpose of $\bm{z}$. The expectation value wrt a random variable $x$
is denoted as $\mathbb{E}_x$. The probability of an event $A$ is denoted as $\mathbb{P}(A)$.
The indicator function is denoted as $\mathbb{I}(\cdot)$ and it evaluates to $1$ if its
argument is true and $0$ otherwise.

@@

The pdf of $\tilde{y}$ can be written using the Law of Total Expectation
\begin{align}
    p(\tilde{y} | H_1) = \mathbb{E}_{h}\left[p(\tilde{y} | h, H_{1})\right]
                 = \int_{-\infty}^{+\infty} p(\tilde{y} | h, H_1)p(h)\;\mathrm{d}h.
\end{align}

Recall that the energy detection rule can be expressed as
\begin{equation}
    d_\delta (\tilde{y}) = \mathbb{I}(\tilde{y} > \delta)
\right.
\end{equation}
in which $\delta$ is a strictly positive real number known as energy threshold,
and $d_\delta (\tilde{y}) = j,~j \in \{0,1\}$, means that the detector has decided
in favor of the hypothesis $H_j$.

As a result, the probabilities of false alarm and miss detection can
be written as
\begin{align}
    p_f &\triangleq \mathbb{P}\left(d_\delta(\tilde{y}) = 1 | H_0\right) = 1 -  p(\delta | H_0),\label{eq:pf} \\
    p_d &\triangleq \mathbb{P}\left(d_\delta(\tilde{y}) = 0 | H_1\right) = \mathbb{E}_{h}\left[p(\delta, h | H_1)\right],
\label{eq:pd}
\end{align}

\subsection{Parameter Estimation in $\alpha$-$\mu$ fading}

The $\alpha$-$\mu$ fading probability density is given as
\begin{align}
p(h) = \dfrac{\alpha \mu^\mu h^{\alpha\mu -1}}{\Gamma(\mu)}\exp\left(-\mu h^\alpha\right),
\end{align}
$h \in \mathbb{R}_{+}$, $\alpha \in \mathbb{R}_{+}$, and $\mu \in \mathbb{R}_{+}$.

Given a vector $\bm{h} \in \mathbb{R}^{n \times 1}$ of iid samples from the
$\alpha-\mu$ distribution, we would like to estimate the value of $\alpha$.
The pdf of $\bm{h}$ is given as
\begin{align}
    p(\bm{h}) = \prod_{i=1}^{n}p(h_i) = \prod_{i=1}^{n}\dfrac{\alpha \mu^\mu h_i^{\alpha\mu -1}}
    {\Gamma(\mu)}\exp\left(-\mu h_i^\alpha\right)\\
    = \left(\dfrac{\alpha \mu^\mu }{\Gamma(\mu)}\right)^{n}
    \exp\left(-\mu\sum_{i=1}^{n}h_i^{\alpha}\right)\prod_{i=1}^{n}h_i^{\alpha\mu -1}
\end{align}
and the log likelihood function (up to an additive constant that does not depend
upon $\alpha$) is given as
\begin{align}
    \log p(\bm{h}) \cong n\log\alpha - \mu\sum_{i=1}^{n}h_i^{\alpha} + \alpha\mu\sum_{i=1}^{n}\log h_i
\end{align}
Noting that
\begin{equation}
    \sum_{i=1}^{n}h_i^{\alpha} \geq \sum_{i=1}^{n} h_i^{\alpha_t} + (\alpha - \alpha_t)\sum_{i=1}^{n}h_i^{\alpha_t}\log h_i
\end{equation}
and therefore
\begin{align}
    \log p(\bm{h}) \leq n\log\alpha - \mu\left(\sum_{i=1}^{n} h_i^{\alpha_t} +
    (\alpha - \alpha_t)\sum_{i=1}^{n}h_i^{\alpha_t}\log h_i\right) + \alpha\mu\sum_{i=1}^{n}\log h_i
\end{align}
